---
title: 'Assignment 1 - Solution'
author name: 'Minhajul Islam Nion'
author ID: 'u3266650'
output:
  html_document: rmdformats::readthedown
editor_options: 
  markdown: 
    wrap: 100
---
***

## Data Description
The following code generates the required dataset for this assignment. It will create a data frame with 155 rows and 13 columns. The variables of this data frame are as following:

- `Date`: Date of the reading (in YYYY-MM-DD format)
- `Temperature`: Temperature reading in Celsius
- `Humidity`: Humidity reading as a percentage
- `Pressure`: Atmospheric pressure in millibars
- `WindSpeed`: Wind speed in kilometers per hour
- `WindDirection`: Wind direction
- `DewPoint`: Dew point temperature in Celsius
- `CloudCover`: Cloud cover as a percentage
- `Precipitation`: Precipitation amount in millimeters
- `Visibility`: Visibility distance in kilometers
- `UVIndex`: UV index reading
- `condition`: The global weather condition over the day
- `Location`: The city of the recorded data

```{r, eval=FALSE}
# Set the seed for reproducibility
set.seed(246)

# Create a sequence of dates for few months in 2024
date_March <- seq(from = as.Date("2024-03-01"), to = as.Date("2024-03-31"), by = 1)
date_May <- seq(from = as.Date("2024-05-01"), to = as.Date("2024-05-31"), by = 1)
date_July <- seq(from = as.Date("2024-07-01"), to = as.Date("2024-07-31"), by = 1)
date_October <- seq(from = as.Date("2024-10-01"), to = as.Date("2024-10-31"), by = 1)
date_December <- seq(from = as.Date("2024-12-01"), to = as.Date("2024-12-31"), by = 1)

dates = c(date_March, date_May, date_July, date_October, date_December)

# Determine the correct number of repetitions for Location
cities <- c("Canberra", "Melbourne", "Sydney")
Location <- rep(cities, length.out = length(dates))

# Create a data frame to store the weather data
weather <- data.frame(
  Date = dates,
  Temperature = round(runif(length(dates), 2, 37), 1),
  Humidity = sample(50:100, length(dates), replace = TRUE),
  Pressure = sample(995:1015, length(dates), replace = TRUE),
  WindSpeed = sample(5:20, length(dates), replace = TRUE),
  WindDirection = sample(c("N", "NE", "E", "SE", "S", "SW", "W", "NW"), length(dates), replace = TRUE),
  DewPoint = round(runif(length(dates), 10, 15), 1),
  CloudCover = sample(0:100, length(dates), replace = TRUE),
  Precipitation = round(runif(length(dates), 0, 15), 1),
  Visibility = sample(5:20, length(dates), replace = TRUE),
  UVIndex = sample(1:12, length(dates), replace = TRUE),
  Condition = sample(c("Sunny", "Partly Cloudy", "Rainy", "Snowy"), length(dates), replace = TRUE),
  Location = Location
)

```

***

## Initial Section:

## Submission for Part A: Data Understanding


```{asis, echo=TRUE}

----------------------------------

1- The data set description goes here

This dataset includes five months' worth of weather data from three cities in 2024: Canberra, Melbourne, and Sydney. The dataset contains characteristics like temperature, humidity, pressure, wind speed, etc., and each row represents a single day. This dataset is useful for researching climate behaviour, examining weather trends, and making predictions for various weather scenarios in certain cities.




----------------------------------
```
```{r}
#----------------------------------#

# 2-  The code for task 2 goes here
# Generate summary statistics for the dataset
summary(weather)



#----------------------------------#

# 3-  The code for task 3 goes here
# Generate a histogram of temperature readings
hist(weather$Temperature, main = "Temperature Distribution", xlab = "Temperature (°C)", col = "lightblue")



#----------------------------------#
```
```{asis, echo=TRUE}
4- The reflection and notes for task 2 and 3 goes here

For task 2, Based on the summary statistics, we can see that the temperature varies from 2°C to 37°C, and the humidity levels vary from 50% to 100%. The pressure stays between 995 and 1015 millibars, which is a somewhat small range.



For task 3, The temperature measurements' histogram, which displays a fairly uniform distribution over the range with some concentration in the 20–30°C range, represents the usual weather for the cities that were chosen.



----------------------------------

```

## Submission for part B: Vector and Matrix Manipulation


```{r}
#----------------------------------#

# 1-  The code for task 1 goes here
# Calculate the average temperature per month
average_temp <- tapply(weather$Temperature, format(weather$Date, "%Y-%m"), mean)
average_temp



#----------------------------------#

# 2-  The code for task 2 goes here
# Calculate the average humidity for each city
average_humidity <- tapply(weather$Humidity, list(weather$Location, format(weather$Date, "%Y-%m")), mean)
average_humidity



#----------------------------------#

# 3-  The code for task 3 goes here
# Create a matrix for average monthly temperature, humidity, pressure, and wind speed
monthly_matrix <- aggregate(cbind(Temperature, Humidity, Pressure, WindSpeed) ~ format(Date, "%Y-%m"), weather, mean)
monthly_matrix



#----------------------------------#

# 4-  The code for task 4 goes here
# Create a matrix for each city's average temperature, humidity, pressure, and wind speed
city_matrix <- aggregate(cbind(Temperature, Humidity, Pressure, WindSpeed) ~ Location, weather, mean)
city_matrix



#----------------------------------#

# 5-  The code for task 5 goes here
# Create an array with monthly and city data
city_array <- array(c(monthly_matrix$Temperature, monthly_matrix$Humidity, monthly_matrix$Pressure, monthly_matrix$WindSpeed),
                    dim = c(5, 3, 4),
                    dimnames = list(month = unique(format(weather$Date, "%Y-%m")),
                                    city = unique(weather$Location),
                                    variables = c("Temp", "Humidity", "Pressure", "WindSpeed")))
city_array


#----------------------------------#

# 6-  The code for task 6 goes here
# Perform matrix multiplication of the transpose with the vector of average temperatures
matrix_product <- t(as.matrix(monthly_matrix[,2:5])) %*% as.vector(average_temp)
matrix_product



#----------------------------------#

# 7-  The code for task 7 goes here
# Custom function to flag above or below average temperatures
compare_temp <- function(temp, avg_temp) {
  ifelse(temp > avg_temp, 'Above', 'Below')
}

# Apply the function and add the results as a new column
weather$Temp_Comparison <- sapply(weather$Temperature, compare_temp, avg_temp = average_temp[format(weather$Date, "%Y-%m")])
head(weather)



#----------------------------------#

```

*** 

## Submission for part C: Looping and Conditional Statements
Please follow this structure:

```{r}
#----------------------------------#

# 1-  The code for task 1 goes here

# Calculate the average pressure for each month using a loop
avg_pressure <- sapply(split(weather$Pressure, format(weather$Date, "%Y-%m")), mean)
avg_pressure


#----------------------------------#

# 2-  The code for task 2 goes here
# Count days with temperature above 25°C in Sydney
days_above_25_sydney <- sum(weather$Temperature > 25 & weather$Location == 'Sydney')
days_above_25_sydney



#----------------------------------#

# 3-  The code for task 3 goes here
# Calculate the average humidity for days with temperature below 21°C in Canberra
avg_humidity_canberra <- mean(weather$Humidity[weather$Temperature < 21 & weather$Location == 'Canberra'])
avg_humidity_canberra



#----------------------------------#

# 4-  The code for task 4 goes here
# Count days with UV Index above 7 in Canberra and Sydney
days_high_uv <- sum(weather$UVIndex > 7 & (weather$Location == 'Canberra' | weather$Location == 'Sydney'))
days_high_uv



#----------------------------------#

# 5-  The code for task 5 goes here
# Create a matrix for monthly total precipitation by city
total_precipitation <- matrix(0, nrow=5, ncol=3, dimnames = list(month = unique(format(weather$Date, "%Y-%m")), city = unique(weather$Location)))
for (i in 1:nrow(total_precipitation)) {
  for (j in 1:ncol(total_precipitation)) {
    total_precipitation[i, j] <- sum(weather$Precipitation[format(weather$Date, "%Y-%m") == rownames(total_precipitation)[i] & weather$Location == colnames(total_precipitation)[j]])
  }
}
total_precipitation



#----------------------------------#
```

*** 

## Submission for part D: Data Frame Manipulation
Please follow this structure:

```{r}
#----------------------------------#

# 1-  The code for task 1 goes here
library(dplyr)
# File names
file_names <- c("201808.csv", "201809.csv", "201810.csv", "201811.csv", "201812.csv")

#----------------------------------#

# 2-  The code for task 2 goes here
# Check the dimensions of the combined data frame
weather_data_list <- lapply(file_names, function(file) {
  read.csv(file, skip = 7)
})

# Combine all data frames into one
weather <- do.call(rbind, weather_data_list)

# Check the dimensions of the combined data frame
print(dim(weather))


#----------------------------------#

# 3-  The code for task 3 goes here
# Check each column's structure and summary
for (col in names(weather)) {
  cat("\nStructure of", col, ":\n")
  str(weather[[col]])
  cat("\nSummary of", col, ":\n")
  print(summary(weather[[col]]))
}


#----------------------------------#

# 4-  The code for task 4 goes here
# Simulate a column with all NAs
weather <- weather[, colSums(!is.na(weather)) > 0]




#----------------------------------#
# 5-  The code for task 5 goes here
# Ensure the Date column is correctly formatted as Date type
colnames(weather) <- gsub(" |\\.", "_", colnames(weather))





#----------------------------------#

# 6-  The code for task 6 goes here
# Create new columns for month and year from the Date column
weather <- weather %>%
  mutate(Date = as.Date(Date, format="%Y-%m-%d"),  # Ensure Date is in Date format
         Month = format(Date, "%m"),
         Year = format(Date, "%Y"))

# Display the first few rows to check
head(weather)



#----------------------------------#

# 7-  The code for task 7 goes here
# Save the combined weather data file with all changes to a single CSV file
write.csv(weather, "combined_weather_data.csv", row.names = FALSE)


#----------------------------------#
```

***

## Overall Conclusion

I gained knowledge of a variety of data manipulation strategies from this assignment, such as looping, matrix operations, data wrangling, and condition handling. My practical abilities in handling real-world data sets were enhanced by working with weather data, and the insights gained from data analysis could be useful in climate studies and weather forecasting.

