---
title: 'Assignment 1 - Solution'
author name: 'provide your name'
author ID: 'provide your student id'
output:
  html_document: rmdformats::readthedown
editor_options: 
  markdown: 
    wrap: 100
---
***

## Data Description
The following code generates the required dataset for this assignment. It will create a data frame with 155 rows and 13 columns. The variables of this data frame are as following:

- `Date`: Date of the reading (in YYYY-MM-DD format)
- `Temperature`: Temperature reading in Celsius
- `Humidity`: Humidity reading as a percentage
- `Pressure`: Atmospheric pressure in millibars
- `WindSpeed`: Wind speed in kilometers per hour
- `WindDirection`: Wind direction
- `DewPoint`: Dew point temperature in Celsius
- `CloudCover`: Cloud cover as a percentage
- `Precipitation`: Precipitation amount in millimeters
- `Visibility`: Visibility distance in kilometers
- `UVIndex`: UV index reading
- `condition`: The global weather condition over the day
- `Location`: The city of the recorded data

```{r, eval=FALSE}
# Set the seed for reproducibility
set.seed(246)

# Create a sequence of dates for few months in 2024
date_March <- seq(from = as.Date("2024-03-01"), to = as.Date("2024-03-31"), by = 1)
date_May <- seq(from = as.Date("2024-05-01"), to = as.Date("2024-05-31"), by = 1)
date_July <- seq(from = as.Date("2024-07-01"), to = as.Date("2024-07-31"), by = 1)
date_October <- seq(from = as.Date("2024-10-01"), to = as.Date("2024-10-31"), by = 1)
date_December <- seq(from = as.Date("2024-12-01"), to = as.Date("2024-12-31"), by = 1)

dates = c(date_March, date_May, date_July, date_October, date_December)

# Determine the correct number of repetitions for Location
cities <- c("Canberra", "Melbourne", "Sydney")
Location <- rep(cities, length.out = length(dates))

# Create a data frame to store the weather data
weather <- data.frame(
  Date = dates,
  Temperature = round(runif(length(dates), 2, 37), 1),
  Humidity = sample(50:100, length(dates), replace = TRUE),
  Pressure = sample(995:1015, length(dates), replace = TRUE),
  WindSpeed = sample(5:20, length(dates), replace = TRUE),
  WindDirection = sample(c("N", "NE", "E", "SE", "S", "SW", "W", "NW"), length(dates), replace = TRUE),
  DewPoint = round(runif(length(dates), 10, 15), 1),
  CloudCover = sample(0:100, length(dates), replace = TRUE),
  Precipitation = round(runif(length(dates), 0, 15), 1),
  Visibility = sample(5:20, length(dates), replace = TRUE),
  UVIndex = sample(1:12, length(dates), replace = TRUE),
  Condition = sample(c("Sunny", "Partly Cloudy", "Rainy", "Snowy"), length(dates), replace = TRUE),
  Location = Location
)

```

***

## Initial Section:

## Submission for Part A: Data Understanding
Please follow this structure:

```{asis, echo=TRUE}

----------------------------------

1- The data set description goes here

This dataset ....



----------------------------------
```

```{r}
#----------------------------------#

# 2-  The code for task 2 goes here



#----------------------------------#

# 3-  The code for task 3 goes here



#----------------------------------#
```
```{asis, echo=TRUE}
4- The reflection and notes for task 2 and 3 goes here

For task 2, ...


For task 3, ...


----------------------------------

```

## Submission for part B: Vector and Matrix Manipulation
Please follow this structure:

```{r}
#----------------------------------#

# 1-  The code for task 1 goes here



#----------------------------------#

# 2-  The code for task 2 goes here



#----------------------------------#

# 3-  The code for task 3 goes here



#----------------------------------#

# 4-  The code for task 4 goes here



#----------------------------------#

# 5-  The code for task 5 goes here



#----------------------------------#

# 6-  The code for task 6 goes here



#----------------------------------#

# 7-  The code for task 7 goes here



#----------------------------------#

```

*** 

## Submission for part C: Looping and Conditional Statements
Please follow this structure:

```{r}
#----------------------------------#

# 1-  The code for task 1 goes here



#----------------------------------#

# 2-  The code for task 2 goes here



#----------------------------------#

# 3-  The code for task 3 goes here



#----------------------------------#

# 4-  The code for task 4 goes here



#----------------------------------#

# 5-  The code for task 5 goes here



#----------------------------------#
```

*** 

## Submission for part D: Data Frame Manipulation
Please follow this structure:

```{r}
#----------------------------------#

# 1-  The code for task 1 goes here



#----------------------------------#

# 2-  The code for task 2 goes here



#----------------------------------#

# 3-  The code for task 3 goes here



#----------------------------------#

# 4-  The code for task 4 goes here



#----------------------------------#
# 5-  The code for task 5 goes here



#----------------------------------#

# 6-  The code for task 6 goes here



#----------------------------------#

# 7-  The code for task 7 goes here



#----------------------------------#
```

***

## Overall Conclusion

Your overall reflection about this assignment goes here ...
